Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-09-16 sections/background_related.tex
Updates from Overleaf

\section{Background and Related Works}

\todo{Sound-matching is an amalgamation of various fields and  requires knowledge of digital signal processing, audio representation, and heuristic search. Here we introduce the various fields and ideas behind sound-matching, before we use these ideas in a historical framing of past sound-matching works.}

\subsection{Digital Audio Synthesis}
\label{sec:dsp}
A digital audio synthesizer is any software used for the creation and manipulation of digital audio. Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds. These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.

The simplest form of DSP function is a sinusoidal tone. For example, consider:
\[ x(n) = sin( 2\pi n T)\]

where $T$ represents time in seconds, and $n$ is the frequency parameter. If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second. Waveform generators are called oscillators for this reason.

Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}. Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.

Particularly in earlier works, frequency/amplitude modulators (\gls{FM}/\gls{AM}) has been the synthesizer of choice in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}. This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}. A common implementation is the use of one carrier, one modulator, and the corresponding envelopes:
\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
\]
Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.

Compared to the number of possible approaches to sound synthesis, a relatively small subset of synthesis methods have been analyzed in \textit{isolation} for sound-matching. Notable methods here are additive and/or subtractive synthesizers~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.

\highlight{
What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable. Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section. 
}

An important recent development in digital audio synthesis is the increase of interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}. DDSP is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP. Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound. These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.

@@

Here we discuss four methods of audio representation and the corresponding loss functions. We also include a technical justification for our novel methods of defining sound similarity using the Scale-Invariant Spectrogram Error (SIMSE) and Dynamic Time Warping (DTW) techniques.

@@

\subsubsection{Fourier Spectrograms}
\label{sec:fourier_specs}
Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain. Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window. They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}. There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.

What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach. For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}. Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.

\textbf{Why use Scale-Invariant Spectrogram for audio loss}
A limitation of L1/L2 distances on spectrograms is their sensitivity to global gain: 
two sounds with identical spectral shape but different amplitude can yield large losses. 
Scale-Invariant Mean Squared Error (SIMSE) addresses this by normalizing the energy of the compared signals, thereby focusing on proportional differences in spectral envelopes. 
This can be advantageous for subtractive synthesis (used by one of our synthesizers), where filter cutoffs reshape the spectral balance which can cause changes in total energy. 
\todo{Perceptually, human listeners are often invariant to global amplitude changes, attending instead to relative spectral shapes}. 
Although SIMSE has been used in other contexts (e.g., image/audio reconstruction), its application as a differentiable spectrogram loss in iterative sound-matching is, to our knowledge, novel.

@@

\subsubsection{Dynamic Envelope Warping}
Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}. Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$. When the series are \textit{warped}, these indices change to expand or contract different portions of the series. To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions. In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.

\textbf{Why use DTW for audio loss?}
DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance. The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}. Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}. DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
By applying DTW specifically to amplitude envelopes, we align with perceptual cues of loudness modulation while maintaining differentiability. 
To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).

\subsection{Automatic Sound Matching}
\label{sec:sound_matching_definition}
The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
\begin{itemize}
    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
    \item $t$: The target sound which we want to replicate or imitate. 
    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
\end{itemize}

\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}. Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.

@@

\subsection{Historical Framing of Sound-Matching}
Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature. This leads us to a more detailed historical analysis of past works in sound-matching.

Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds using the simple FM synthesizer described in Section~\ref{sec:dsp}. The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged. Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.

Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching. Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}. They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer. The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results. As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.

Recent years have seen more works in sound-matching using supervised machine learning techniques. In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test). The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models. For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function. For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline. Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.

Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}. Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search. This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter). A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss. After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}. This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss. Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}. Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.

@@

During the training phase, this metric is calculated alongside the neural parameter estimator. After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.

@@

Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
Finetuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}. Barkan \textit{et al.} report that the application of ITF generally worsened the spectral and manual performance metrics, perhaps due to the discrepancy between the true and approximated synthesizer or overfitting.

@@

\subsection{What is Lacking In the Field}
\label{sec:lacking}
%-----> much of the original text (available in older version) is removed for brevity)

Having looked at the current available literature, we note four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
\begin{enumerate}
    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output? 
    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching. The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently. In differentiable settings, this problem causes gradient descent updates to local minima. 
    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
\end{enumerate}

 % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods. We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer. This hypothesis targets the issues of \LossSelect~and \SynthSelect.

  % Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers. What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests. \textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}. If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.

>>>>

\section{Background and Related Work}
\label{sec:background_related}

Sound-matching sits at the intersection of digital signal processing, audio representation, and optimization. 
In this section we first formalize the sound-matching task, then review core synthesis methods, similarity measures, and optimization approaches. 
We conclude with a survey of related work and a discussion of gaps in the field.

\subsection{Formalization of Sound-Matching}
\label{sec:sound_matching_definition}
Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure. 
The key components are:
\begin{itemize}
    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
    \item $x$: The synthesizer output, $x = g(\theta)$.
    \item $t$: The target sound to be replicated or imitated. 
    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
    \item $L$: A loss function that measures distance between $x$ and $t$, typically via $L(\theta,t) = d(\phi(x), \phi(t))$ for some metric $d$.
\end{itemize}

This formalization highlights that sound-matching is not a single well-defined optimization problem, but rather depends on design choices for $g$, $\phi$, and $L$. 
Moreover, the ``correct'' solution may not exist in a strict sense: depending on the artistic goal, multiple parameter settings may yield acceptable or even preferable outputs.  In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.

\subsection{Digital Signal Processing and Synthesis}
\label{sec:dsp}

A digital audio synthesizer generates or processes audio by chaining digital signal processing (\gls{DSP}) functions. 
Each function is parameterized, and the set of parameters defining a chain of DSP functions is called a synthesizer program.

The simplest DSP function is a sinusoidal oscillator: 
\[
x[n] = \sin\!\left(2 \pi f \frac{n}{SR}\right),
\]
where $f$ is frequency in hertz, $SR$ is the sampling rate, and $n$ is the discrete time index. 
At a sampling rate of $SR$ samples per second, a 1\,Hz oscillator completes one cycle per second. 
Such computations form the foundation of digital audio synthesis~\cite{lyons1997understanding}.

Since the advent of DSP in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric functions have been proposed, including oscillators, filters, equalizers, and envelopes~\cite{lyons1997understanding,russ1999sound}. 
Sound design involves modifying these parameters until a desired output is reached~\cite{roads1996computer,pinch2004analog}.

Early sound-matching research focused heavily on frequency and amplitude modulation (\gls{FM}/\gls{AM}) synthesis~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}, which is simple to implement yet expressive~\cite{chowning1973synthesis}. 
Other approaches studied in isolation include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.

\highlight{By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable. 
Recent work has also explored commercial Virtual Studio Technology (VST) synthesizers~\cite{yee2018automatic,esling2019flow}. 
While powerful, such systems obscure the interactions between modules, losses, and outputs. 
In contrast, small-scale differentiable synthesizers built from isolated DSP blocks offer interpretability, making them particularly useful for benchmarking. }

Finally, recent years have seen a growing interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}, which integrates gradient-based optimization~\cite{goodfellow2016deep,boyd2004convex} with DSP building blocks. 
Implementing complex DSP functions in a differentiable manner remains challenging, and robust differentiable audio similarity measures are still under active investigation~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.

@@

Here we discuss four methods of audio representation and the corresponding loss functions, including technical justifications for our novel methods.

@@

\subsubsection{Fourier Spectrograms}
\label{sec:fourier_specs}
Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain. Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window. They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}. There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.  
What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach. For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}. Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.

A common limitation of spectrogram-based losses is their sensitivity to global gain: two signals with identical spectral shape but different overall amplitude can yield large errors under L1/L2 norms. 
Here we propose and test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1/L2 norms in spectrogram comparisons. SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes. 
Perceptually, listeners are often tolerant of loudness changes while remaining sensitive to timbral shape, therefore this property could be advantageous for subtractive synthesis, where filter cutoffs reshape spectral balance without predictable changes in total energy. Although SIMSE has been applied in other contexts such as audio reconstruction, to our knowledge its use as a differentiable spectrogram loss in iterative sound-matching is novel.

@@

\subsubsection{Dynamic Envelope Warping}
Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}. Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$. When the series are \textit{warped}, these indices change to expand or contract different portions of the series. To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions. In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}. DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance, such that the distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}. Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.

DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
By applying DTW specifically to amplitude envelopes, we can match two sounds with identical rates ofloudness modulation regardless of alignments. 
To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).

@@

\subsection{Historical Framing of Sound-Matching}
Given the attributes discussed above, Table~\ref{tab:summary} summarizes relevant literature. Here we expand with a historical analysis of past sound-matching work.

Perhaps the earliest foundational study is Justice~\cite{justice1979analytic}, who analytically decomposed and recreated sounds using a simple FM synthesizer (see Section~\ref{sec:dsp}). 
Subsequent work retained this FM structure but explored new heuristics for parameter search. 
For example, Horner \textit{et al.}~\cite{horner1993machine} applied genetic algorithms (GAs) to resynthesize sounds with one modulator and one carrier, measuring similarity via the McAulay–Quatieri method~\cite{mcaulay1986speech}.

Later studies introduced more complex synthesis models, such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter}. 
Mitchell and Creasy~\cite{mitchell2007evolutionary} highlighted the difficulty of disentangling synthesizer limitations from optimization inefficiency. 
They proposed a \textit{contrived methodology} in which the best heuristic for in-domain FM resynthesis should also generalize to out-of-domain targets (e.g., muted trumpet tones~\cite{opolko1989mcgill}). 
However, limited testing produced contradictory results, and they concluded that changes to the synthesizer, loss, or sound domain effectively redefine the search space~\cite{mitchell2007evolutionary}.

Recent years have seen more works in sound-matching using supervised machine learning techniques. In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test). The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models. For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function. For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline. Importantly, informal listening tests found the results unsatisfactory, likely due to the synthesizer’s 155-parameter complexity.

Masuda \textit{et al.} also applied supervised learning~\cite{masuda2021soundmatch}. Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search. This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter) alongside a NN model pre-trained with P-Loss on an in-domain dataset of randomly selected parameters. After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}. This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss. Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}. Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.

@@

This metric is calculated alongside the neural parameter estimator. Once trained, PNP enabled the fast differentiable optimization of computationally expensive loss functions such as JTFS with FM and physical models~\cite{han2023perceptual,han2024learning}.

@@

Barkan \textit{et al.} attempted to  improve the IS2 model with Inference-Time
Finetuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}. Barkan \textit{et al.} report that the application of ITF generally worsened the spectral and manual performance metrics, perhaps due to the discrepancy between the true and approximated synthesizer or overfitting.

@@

% \subsection{What is Lacking In the Field}
% \label{sec:lacking}
% %-----> much of the original text (available in older version) is removed for brevity)

% Having looked at the current available literature, we note four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
% \begin{enumerate}
%     \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output? The few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods\todo{cite?}.
%     \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching. The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
%     \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}. In differentiable settings, this problem causes gradient descent updates to local minima. 
%     \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
% \end{enumerate}

% \subsection{What Approach Should We Take?}
% We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems. To address the problems of \LossSelect{} and \SynthSelect{}, we pursue a methodology with 3 key characteristics that do not appear together in other works.

% 1. Simple differentiable synthesizers using common DSP functions, filters, and oscillators would be beneficial in addressing the \SynthSelect{} problem. Isolating small-scale, low-parameter synthesizers avoids the confounding effects of large neural proxies or embedding models, making it possible to directly analyze the interaction between synthesis structure and loss function.

% 2. Evaluation of experiments under multiple loss functions: Addressing the \LossSelect{} problem requires that multiple loss functions are tested. A comparison of performance of different loss functions is rare in previous works~\todo{cite}.

% 3. Parameters optimized directly with gradient descent without additional neural network approximations. This would complement prior neural and proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.

% These characteristics would fill a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.

% % 4. Past synths are either simple differentiable functions (commonly FM), or complex VSTs where we can't isolate the interactions between loss and functions, here we use isolated building blocks of DSP functions in a differentiable environment.

%  % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods. We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer. This hypothesis targets the issues of \LossSelect~and \SynthSelect.
